k-plus Anticlustering: An Improved k-means Criterion for Maximizing Between-Group Similarity

### 2024-02-21 paper.Rmd
minor

- Start with use case: Stimulus assignment in within-subjects experiments (and stick with it throughout the paper, only briefly mention other use cases)
  * Needed: high similarity between stimulus set to reduce variance in responses attributed to other factors than the experimental manipulation. This is possible using `anticlust`, which has been applied many times since its introduction in @papenberg2020 (cite references)
  * Also needed (point of this paper!): High pairwise dissimilarity of stimuli in the same set. 
- Simultaneous optimization of between-set similarity and pairwise within-set dissimilarity was introduced by Brusco et al.'s alogrithm, which is also available in `anticlust` since 2021 (version 0.6.0)
- Here, we provide researchers with a practical introduction to their method and also present several extensions, which allow it to cover more use cases
  (1) Always return optimal solution
  (2) Use different distance matrix for dispersion and diversity (if different information is needed for both criteria, or to induce custom cannot-link constraints)
  (3) Extend the BILS's capacities towards other measures of between-group similarity (k-means / k-plus)
- Technical introduction to "vanilla" BILS
- Technical introduction to our extensions
  * Optimal dispersion method: Basic algorithm + graph coloring ILP
    * How to maintain optimal dispersion while the BILS runs? Outline three approaches!
  * Different distance matrices: does not change algorithm, just computation / implementation
  * Extend towards k-means: Use squared Euclidean distance (not really extension, just different input); compute "average" diversity if group sizes are unequal (k-plus is basically the same, but do not use the additional k-plus variables for the dispersion distance matrix)
- Simulation Study 1: Two research questions
  (1) How often does vanilla BILS find optimal solutions
  (2) Which of the three extension is best for optimizing between-group similarity while maintaining optimal dispersion?
- Simulation Study 2: More difficult conditions (do not continuous data; K > 4 groups)
- Application (presenting R code)
  * Optimize k-plus criterion using norming data; dispersion is based on pairwise orthographic dissimilarity between words
- Discussion

>>>>

- Start with use case: Stimulus assignment in within-subjects experiments (and stick with it throughout the paper, only briefly mention other use cases)
  * Needed: high similarity between stimulus set to reduce variance in responses attributed to other factors than the experimental manipulation. This is possible using `anticlust`, which has been applied many times since its introduction in @papenberg2020 (cite references)
  * Also needed (point of this paper!): High pairwise dissimilarity of stimuli in the same set. 
- Simultaneous optimization of between-set similarity and pairwise within-set dissimilarity was introduced by Brusco et al.'s alogrithm, which is also available in `anticlust` since 2021 (version 0.6.0)
- Here, we provide researchers with a practical introduction to their method and also present several extensions, which allow it to cover more use cases
  (1) Always return optimal solution
  (2) Use different distance matrix for dispersion and diversity (if different information is needed for both criteria, or to induce custom cannot-link constraints)
  (3) Extend the BILS's capacities towards other measures of between-group similarity (k-means / k-plus)
- Technical introduction to "vanilla" BILS
- Technical introduction to our extensions
  * Optimal dispersion method: Basic algorithm + graph coloring ILP
    * How to maintain optimal dispersion while the BILS runs? Outline three approaches!
  * Different distance matrices: does not change algorithm, just computation / implementation
  * Extend towards k-means: Use squared Euclidean distance (not really extension, just different input); compute "average" diversity if group sizes are unequal (k-plus is basically the same, but do not use the additional k-plus variables for the dispersion distance matrix)
- Simulation Study
  (1) How often does vanilla BILS find optimal solutions
  (2) Which of the three extension is best for optimizing between-group similarity while maintaining optimal dispersion?
- Application (presenting R code)
  * Optimize k-plus criterion using norming data; dispersion is based on pairwise orthographic dissimilarity between words
- Discussion
